---
title: "Intro to igraph"
author: '@plzbeemyfriend'
date: "March 10, 2018"
output: 
    ioslides_presentation:
        widescreen: true
        smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(igraph)
library(diagram)
```

## R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Why Graph - Data schema is the actual data model

## Why Graph - Graph databases scale really well

## Why Graph - Recommendation Systems

## Why Graph - Recommendation Systems Continued

## Why Graph - Recommendation Systems Continued Again

## Terminology - Undirected Graph
- **A**, **B**, and **C** are nodes/vertices 
- **1** and **2** are edges/links

```{r}
my_edge_length <- .09
my_text_size <- 1.3
openplotmat()
pos <- coordinates(c(3, 3, 3))
arrow_1 <- straightarrow(from = pos[1,], to = pos[2,], arr.width = 0)
arrow_2 <- straightarrow(from = pos[2,], to = pos[3,], arr.width = 0)
textellipse(mid = pos[1,], radx = my_edge_length, lab = "A", box.col = 'lightgrey', cex = 1.5)
textellipse(mid = pos[2,], radx = my_edge_length, lab = 'B', box.col = 'lightblue', cex = 1.5)
textellipse(mid = pos[3,], radx = my_edge_length, lab = 'C', box.col = 'lightgrey', cex = 1.5)
text(arrow_1[1], arrow_2[2] + .1, "1", cex = 1.5)
text(arrow_2[1], arrow_2[2] + .1, '2', cex = 1.5)
```

## Terminology - Directed Graph
- **A**, **B**, and **C** are nodes/vertices 
- **1** and **2** are edges/links
- Edge **1** is coming **IN** to node **B**
- Edge **2** is going **OUT** of node **B**
- Node **B** is at the **HEAD** of edge **1**
- Node **B** is at the **TAIL** of edge **2**

```{r terminology}
my_edge_length <- .09
my_text_size <- 1.3
openplotmat()
pos <- coordinates(c(3, 3, 3))
arrow_1 <- straightarrow(from = pos[1,], to = pos[2,])
arrow_2 <- straightarrow(from = pos[2,], to = pos[3,])
textellipse(mid = pos[1,], radx = my_edge_length, lab = "A", box.col = 'lightgrey', cex = 1.5)
textellipse(mid = pos[2,], radx = my_edge_length, lab = 'B', box.col = 'lightblue', cex = 1.5)
textellipse(mid = pos[3,], radx = my_edge_length, lab = 'C', box.col = 'lightgrey', cex = 1.5)
text(arrow_1[1], arrow_2[2] + .1, "1", cex = 1.5)
text(arrow_2[1], arrow_2[2] + .1, '2', cex = 1.5)
```
## Terminology - Simple Graph
- unweighted
- undirected
- no loops
- no multiple edges

## Terminology - Multigraph
- None of the previous restrictions

## Terminology Continued
- A **walk** is a sequence of vertices and edges
- A **path** is a walk with no repeating vertices with the possible exception of the first and final vertices.

## Creating graphs from scratch
1. `make_empty_graph()`: Initialize and empty graph to build on
2. `vertices()`: Create the vertices/nodes of the graph
3. `edges()`: Create the edges/links connecting the vertices

## Create an undirected graph from scratch
```{r undirected_scratch, echo = T}
library(igraph)

g_undirected <- make_empty_graph(directed = F) +
  vertices(c('a', 'b', 'c', 'd', 'e'),
           type = "letters",
           order = 1, 2, 3, 4, 5) +
  edges(c('a', 'b',
          'b', 'c',
          'b', 'd',
          'c', 'e'),
        type = "connection",
        weight = c(1, 1, 2, 2))
```

## Information from undirected graph
```{r info_undirected_scratch, echo = T}
g_undirected
```
```{r plot_undirected_scratch, echo = F}
set.seed(1234)
plot(g_undirected,
     layout = layout_nicely,
     edge.arrow.size = .3)
```

## Create a directed graph from scratch
```{r directed_scratch, echo = TRUE}
g_directed <- make_empty_graph(directed = T) +
  vertices(c('a', 'b', 'c', 'd', 'e'),
           type = "letter",
           order = c(1, 2, 3, 4, 5)) +
  edges(c('a', 'b',
          'b', 'c',
          'b', 'd',
          'c', 'e'),
        type = "connection",
        weight = c(1, 1, 2, 2))
```

## Information from directed graph
```{r info_directed_scratch, echo = T}
g_directed
```
```{r plot_directed_scratch, echo = F}
set.seed(1234)
plot(g_directed,
     edge.curved = F,
     layout = layout_nicely,
     edge.arrow.size = .3)
```

## Creating graphs from data frames
1. Create and edge list that represents the links of the graph
    + OPTIONAL: Create a node list with information about each unique node
2. `graph_from_data_frame()`: Creates a graph from the edge list

## Creating a graph from a data frame
```{r directed_df, echo=T}
node_list <- data.frame(
  name = c('a', 'b', 'c', 'd', 'e'),
  type = "letter",
  order = c(1, 2, 3, 4, 5)
)
edges(c('a', 'b',
          'b', 'c',
          'b', 'd',
          'c', 'e'),
        type = "connection",
        weight = c(1, 1, 2, 2))
edge_list <- data.frame(
  source = c('a', 'b', 'b', 'c'),
  target = c('b', 'c', 'd', 'e'),
  type = 'connection',
  weight = c(1, 1, 2, 2)
)
g <- graph_from_data_frame(d = edge_list, 
                           directed = T,
                           vertices = node_list)
```

## Information from directed graph
```{r info_directed_df, echo =T}
g
```
```{r plot_directed_df, echo = F}
set.seed(1234)
plot(g,
     edge.curved = F,
     layout = layout_nicely,
     edge.arrow.size = .3)
```

## Accessing/Manipulating Data from graph 
given graph **g**

- `V(g)`: provides a vector of all vertices/nodes of a graph
- `E(g)`: provides a vector of all edges/links of a graph
- `vcount(g)`: provides the number of vertices/nodes of a graph
- `ecount(g)`: provides the number of edges/links of a graph
- `vertex_attr_names(g)`: provides the names of all the attributes stored in vertices/nodes
- `edge_attr_names(g)`: provides the names of all the attributes stored in edges/links

## V()
Vertices can be accessed by their **id**. If the graph is a named graph, then the **name** also functions as an id.
```{r vertex_access, echo = T}
V(g)
V(g)[3]
V(g)['c']
```

## V() Continued
Accessing a vertex using double square bracket notation gives you all the information of the vertex.
```{r vertex_access2, echo = T}
V(g)[['c']]
V(g)[['c']]$order
V(g)['c']$order <- 29
V(g)[['c']]
```

## V() Continued Again
You can access multiple vertices either by providing a vector of names or ids or by **querying the attributes** of the vertex vector. 
```{r vertex_access3, echo = T}
V(g)[[c('a', 'e')]]
V(g)[[order >= 4]]
```

## E()
Edges can have names and can be accessed directly using either a name or an id, however, as graphs get larger, the sheer number and variety of edges make it very impractical to do so. 

```{r, edge_access, echo = T}
E(g)
E(g)[1]
E(g)[[1]]
```

## E() Continued
Instead edges should be accessed by querying their attributes
```{r, edge_access2, echo = T}
E(g)[weight == 2]
E(g)[[weight == 2]]
E(g)[[weight == 2]][1]$weight <- -20000
E(g)[[weight <= 0]]
```

## E() Continued Again
We can also **query edges for the nodes they connect.**
```{r, edge_access3, echo = T}
E(g)['a' %--% 'b']
E(g)[['c' %<-% 'b']]
E(g)[['b' %->% V(g)]]
```

## Subgraphs
Sometimes you may only want to work with a part of the entire graph.

- `induced_subgraph()`: creates a subgraph that includes only the vertices provided and keeps all the edges between those vertices.
- `subgraph.edges()`: creates a subgraph that includes only the edges provided and keeps all the vertices they connect to. Unconnected vertices can also be kept if `delete.vertices = F`

```{r subgraphs, echo = T}
abc_sub <- induced_subgraph(g, c('a', 'b', 'c'))
abc_sub
b_sub <- subgraph.edges(g, E(g)["b" %--% V(g)])
b_sub
```

## Subgraphs Gotchas
In igraph, subgraphs are completely new graphs and their components are completely different from those of the graphs they were created from. The below won't work:

```{r subgraph_gotcha, eval=F, echo = T}
E(g)["b" %--% V(b_sub)]
```

Instead, you will need to use attributes that you know are in both graphs, like names:

```{r subgraph_gotcha_good, echo = T}
E(g)["b" %--% V(b_sub)$name]
```

## Random Graph Generation - Erdos Renyi Game
```{r ERG, echo =T}
set.seed(1234)
g <- erdos.renyi.game(10, .5, directed = T)
V(g)$name <- letters[1:10]
V(g)$randomValue <- runif(vcount(g), 0, 100)
E(g)$randomWeight1 <- runif(ecount(g), 0, 100)
E(g)$randomWeight2 <- runif(ecount(g), 0, 100)
plot(g,
     edge.curved = T)
```

## Degree
```{r echo = T}
degree(g) %>%
  sort(T)

degree(g, mode = 'in') %>%
  sort(T)

degree(g, c('b', 'e')) %>%
  sort(T)
```

## Strength - Weighted Degree
```{r echo=T}
strength(g, mode = 'out', weights = E(g)$randomWeight1)
strength(g, mode = 'in', weights = E(g)$randomWeight1)
strength(g, c('b', 'e'), mode = 'in', weights = E(g)$randomWeight1)
```

## Betweeness
```{r echo = T}
betweenness(g) %>%
  sort(T)
page_rank(g)$vector %>%
  sort(T)
eigen_centrality(g, directed = T)$vector %>%
  sort(T)
```

## Ego

The ego or neighborhood of a vertex V is a vector of vertices that are within a given order or V:
```{r echo = T}
ego(g, order = 1, nodes = 'b', mode = 'all')
ego(g, order = 2, nodes = c('b', 'e'), mode = 'in')
```

## Shortest Paths
```{r echo = T}
shortest_paths(g, 'h', 'c', mode = 'out', output = 'both', weight = E(g)$randomWeight1)$epath[[1]] %>% .[[1:length(.)]]
shortest_paths(g, 'h', 'c', mode = 'out', output = 'both', weight = E(g)$randomWeight2)$epath[[1]] %>% .[[1:length(.)]]
```

## Combo

We can use V(), ego() and induced_subgraph() to get a graph that only includes the nodes of interest. Let's find a Story that does not include Jack as one of the characters.
```{r,  echo = T}
gt75_nodes <- V(g)[[randomValue >= 75]]
gt75_ego <- ego(g, 1, gt75_nodes, 'in')
gt75_ego
```

## Combo Continued
```{r, echo = T}
unlist(gt75_ego)

names_gt75 <- names(unlist(gt75_ego))

induced_subgraph(g, names_gt75)
```


## Pulling edges of interest
```{r eoi, echo = T}
E(g)[randomWeight1 >= 90]
E(g)[randomWeight1 >= 90] %>%
  .['a' %--% V(g)]
```

## Putting edge info into nodes
```{r e2n, echo = T}
V(g)$in_weights <- strength(g, mode = 'in', weights = E(g)$randomWeight1)
V(g)$out_weights <- strength(g, mode = 'out', weights = E(g)$randomWeight1)
V(g)[[1:vcount(g)]]
```

## Putting node info into edges
```{r n2e, echo =T}
E(g)$out_impact <- E(g)$randomWeight1/head_of(g, E(g))$in_weights
E(g)$in_impact <- E(g)$randomWeight1/tail_of(g, E(g))$out_weights
E(g)[[1:10]]
```
## magrittr: an aside
```{r rev_n, echo = T}
rev_n <- function(nodes, func, graph, ...){
  arg <- formals(func) %>%
    names() %>%
    .[sapply(., function(x) grepl('^v|^n', x, T))] %>%
    .[1]
  
  temp <- list()
  temp[[arg]] <- nodes
  temp[['graph']] <- graph
  
  temp <- c(temp, as.list(match.call(expand.dots = F)[["..."]]))
  do.call(func, temp)
}

```

## magrittr: another aside
```{r rev_e, echo = T}
rev_e <- function(edges, func, graph, ...){
  arg <- formals(func) %>%
    names() %>%
    .[sapply(., function(x) grepl('^e', x, T))] %>%
    .[1]
  
  temp <- list()
  temp[[arg]] <- edges
  temp[['graph']] <- graph
  
  temp <- c(temp, as.list(match.call(expand.dots = F)[["..."]]))
  do.call(func, temp)
}
```
## USING THE HELPERS
```{r echo = T}
g %>% 
  E() %>% 
  .[randomWeight1 == max(.$randomWeight1)] %>% 
  rev_e(ends, g) %>% 
  c() %>% 
  rev_n(ego, g, order = 1, 'out') %>% 
  unlist() %>% 
  rev_n(induced.subgraph, g)
```